\documentclass[11pt]{article}

\usepackage{geometry}
\geometry{
  a5paper,
  landscape,
  nohead,
  nofoot,
  vmargin = 15mm,
  hmargin = 20mm,
}

\usepackage[normalem]{ulem}
\usepackage{multicol}
% \usepackage{xcolor}

\usepackage{fontspec}
\defaultfontfeatures{ Ligatures={TeX}, Path = {../fonts/}, }

\setmainfont[
  SmallCapsFeatures = { LetterSpace = 0.2 },
  ItalicFont = Crimson-Italic.otf,
  BoldFont = Crimson-Bold.otf,
  BoldItalicFont = Crimson-BoldItalic.otf,
]{Crimson-Roman.otf}

% \newfontfamily\spectralScFont{SpectralSC-Medium.otf}

\setlength{\parskip}{8pt}
\setlength{\parindent}{0pt}

\newcommand\subSectionFmt[1]{%
  {%
  \normalfont%
  \bfseries\itshape%
  \uline{#1}%
  }%
}

\pagestyle{empty}

\setlength{\columnsep}{20pt}
\setlength{\columnseprule}{0pt}

\begin{document}

% \raggedright

{\LARGE Class Rules}

\begin{multicols}{2}

\subSectionFmt{Kuṭi-Work Before Vinaya Class:}

Sign up for presenting rules and prepare.

There will be a short test before the introduction, with the question:

`What are the ten reasons for the establishing of the Pāṭimokkha?' (Memorize from Pr~1 story or AN~10.31)

\subSectionFmt{Class Materials:}

No phones, tablets or laptops in the class room. Bring pen and paper.

Everyone will get a copy of the general class notes, and a small Pāṭimokkha book. There will be reference books in the room for fact-checking, but use them as little as possible.

\columnbreak

\subSectionFmt{Routine:}

(1)~Opening Pāli phrases. (2)~Origin stories.\\
(3)~Rules. (4)~Quiz and discussion of solutions. (5)~Closing Pāli phrases.

\subSectionFmt{Presenting a Group of Rules:}

20-30 mins, be concise. Focus on the main points.
Aim to speak from memory, prepare short notes as a guide.
Add new quiz questions if you wish.

Arrange with someone to present the origin story, recommended ones are marked with an asterisk on the schedule.

In a discussion, questions not relevant to the topic should be deferred to another occasion.

\subSectionFmt{Presenting Origin Stories:}

Tell us what happens in the story in your own words, focusing on the aspects relevant to the rule. Make a few bullet points to remember, don't read the story from the book.

\end{multicols}

\end{document}
