\chapter{3. Sexual Conduct}
\renewcommand*{\theChapterTitle}{3. Sexual Conduct}

\begin{exam}{\autoExamName}

\begin{problem}

  A bhikkhu is staying at the apartment of a friend, where they organize a small
  gathering, and they start drinking alcohol. The bhikkhu gets drunk, and
  eventually he goes to bed in his room. He wakes up, and finds a woman's
  underwear in his bed, with a note saying `love and kisses', plus a used
  condom.

  Is the bhikkhu pārājika? 

  \bigskip
 
  \begin{answers}{1}
    \bChoices
    \Ans0 No, because he was drunk\eAns
    \Ans0 No, if he was practising trantric freedom and compassion\eAns
    \Ans1 Yes, since there is clear evidence of intercourse\eAns
    \Ans0 Yes, even if he can't remember anything\eAns
    \eChoices
  \end{answers}

  \bigskip

  \textbf{Discussion:} What if he convinces himself that he is pārājika, but
  later finds out that they had played a prank on him?
  
\end{problem}

\problemDivide

\begin{problem}

  Mark the factors which, under \textit{Sg 1}, commit a \textit{thullacāya} offence.

  \begin{answers}{6}
    \bChoices
    \Ans0 object\eAns
    \Ans0 perception\eAns
    \Ans1 intention\eAns
    \Ans1 effort\eAns
    \Ans0 result\eAns
    \eChoices
  \end{answers}

  \bigskip

  \textbf{Discussion:} describe such a situation.

\end{problem}

\problemDivide

\begin{problem*}

  Are there offences?

\begin{parts}

\item
  A women asks to speak with a bhikkhu.
  It is a hot day and she is dressed quite openly.
  For the rest of the day, he continues fantasising about her.

  \bigskip

  \begin{answers}{6}
    \bChoices
    \Ans0 pārājika\eAns
    \Ans0 saṅghādisesa\eAns
    \Ans0 thullacāya\eAns
    \Ans0 pācittiya\eAns
    \Ans0 dukkaṭa\eAns
    \Ans1 no offences\eAns
    \eChoices
  \end{answers}

  \bigskip

  \item Later, the bhikkhu recollects the meeting, starts rubbing himself, and causes an emission.

  \bigskip

  \begin{answers}{6}
    \bChoices
    \Ans0 pārājika\eAns
    \Ans1 saṅghādisesa\eAns
    \Ans0 thullacāya\eAns
    \Ans0 pācittiya\eAns
    \Ans0 dukkaṭa\eAns
    \Ans0 no offences\eAns
    \eChoices
  \end{answers}

  \bigskip

  \textbf{Discussion:} asking women to cover themselves when they come to a
  meeting in what is a normal dress for them.

  \bigskip

\end{parts}

\end{problem*}

\end{exam}
