\chapter{8.A. Robes 1}
\renewcommand*{\theChapterTitle}{8.A. Robes 1}

What is the Pali name of a bhikkhu's upper, lower, and outer robe?

\bigskip

A bhikkhu discovers that the seams of his cotton jacket under the arm-pit where
the cloth was joined, have come apart. What should he do?

\bigskip

Supporters wish to offer robe-cloth to the Community. They bring a piece of cloth as a sample, which
is a white nylon material. Is there an offense in asking them to offer a better
material?

\bigskip

After the Pavarana ceremony, the community holds a Kathina celebration.
At the end, they relinquish the Kathina privileges.
One of the bhikkhus, who didn't really want to relinquish the privileges,
goes on tudong without taking his \emph{saṅghāṭi} with him.
Are there any offenses?

\bigskip

A bhikkhu wants to go tudong without his saṅghāṭi, and asks the community for
permission to do so. Is this allowed?

\bigskip

Is a bhikkhu allowed to travel home without taking his saṅghāṭi? Can he stay one
night at a hospital without it?

\begin{solution}
  He must take his three robes with him when travelling.

  In exceptional cases (short hospital stay), NP~2 allows the community to give
  him permission to be without his robes.
\end{solution}

\bigskip

A bhikkhu receives a nice leather-belt from a friend. Is it allowable?

\bigskip

A bhikkhu embroiders the sign of the Eye of Horus on his meditation blanket. Is it
allowable?

\bigskip

A bhikkhu keeps his three robes in his kuti where he spends the night.
Waking up early while it is still dark, he goes for a walk outside the monastery to watch the Sun rise.
Is there any offense?

\bigskip

A bhikkhu takes some cloth from the stores to his kuti to make a sitting cloth. He
forgets about it for a few weeks. Is there an offense?

\bigskip

A monk is visiting home. His old friends invite him to the skate park. He puts
on a pair of jeans and a black T-shirt to go and see if he can still do an
ollie. Is there an offense?

\bigskip

A bhikkhu asks his mother to buy him a new robe made of silk when she is travelling in Thailand,
even though his mother has asked him not to ask for any more new robes. Is there an offense?

\bigskip

A bhikkhu is chosen by the community to receive the Kathina-robe. What are the
eight Kathina duties? What is procedure when receiving the Kathina robe? What
are the Kathina privileges?

\bigskip

A bhikkhu is travelling by plane. He packs his saṅghāṭi in the hold luggage. After
landing, his hold luggage is missing. He registers the missing luggage with the
airport services, but has to leave without it. The airport delivers his luggage
in a few days. What are his duties?

\bigskip

A bhikkhu wants to mark his robe. He has an ink bottle, and plucks a blade of
grass to make a mark on the robe. Are there offenses?

\bigskip

A monk realises his robe is bigger than the standard measurement 2.25m x 1.5m –
could he confess this to another monk in the monastery?

\bigskip

How should one treat one's robes? If they are torn, or lost or are laid aside, how should one deal with it?
