\chapter{19.A. Women 2}
\renewcommand*{\theChapterTitle}{19.A. Women 2}

\section*{Discussion}

% Ay 1 sitting privately with a woman
% Ay 2 sitting out of earshot with a woman

A bhikkhu is accused of a Sanghadisesa offence. Who decides? Does this principle hold for all offences?

\begin{solution}
  Ultimately the accused Bhikkhu decides and this holds for all rules. 
\end{solution}

\bigskip

What can a community do if a Bhikkhu is considered to have committed an offense but will not admit it?

\begin{solution}
  A community may make a formal act of banishment to that Bhikkhu, so he may no
  longer live with that community.
\end{solution}

\bigskip

% Bhikkhunis Summary of related rules: NP 4-5, NP 17, Pc 21-30, Pd 1-2

You are travelling by car to a teaching engagement and a Siladhara comes along. Is there any offence?

\begin{solution}
  Only if there was prior arrangement.
\end{solution}

\bigskip

Who is a relation? In theory, and in practice?

\begin{solution}
  In theory, going back through seven generations. Practically, are you aware of
  blood relations, and would they feel free to disagree with your requests.
\end{solution}

\bigskip

What is the procedure that is encouraged for bhikkhus to follow at the monasteries of this tradition 
when corresponding with the siladhara?

\begin{solution}
  Not to meet to talk in a private or secluded place,
  having another male around or sitting in an open, public place.

  Male community members are advised not to talk about private topics with them.
\end{solution}

\bigskip

A siladhara wishes to give a gift to a bhikkhu. What is the procedure she should follow?

\bigskip

Does this procedure also apply if you wish to give a gift to a siladhara?

\begin{solution}
  It is better to exchange rather than give gifts.
  A gift should only be made in public, so it is not a special, secret act.

  It is better to suggest that they offer the gift for the community of bhikkhus or/and siladharas,
  rather than for a specific person.
\end{solution}

\bigskip

A siladhara offers to clean a bhikkhu's boots. How should the bhikkhu respond?

\begin{solution}
  No, thank you.
\end{solution}
