\chapter{23.A. Sekhiyas 2}
\renewcommand*{\theChapterTitle}{23.A. Sekhiyas 2}

\begin{exam}{\autoExamName}

  \begin{problem*}
    \textbf{Do} or do \textbf{Not}?

    \bigskip

    \begin{parts}

    \item \TF{D} While eating, a bhikkhu asks for water. Someone hands over a
      water bottle, but he doesn't receive it until he washed his hands.

    \bigskip

    \item \TF{N} Receiving alms-food, a bhikkhu asks a man to replace the tuna
      in tomato with tuna in oil.

    \bigskip

    \item \TF{N} A bhikkhu is explaining an important point about \emph{Nibbāna}
    to a lay person, who is holding a phone in their hands and scrolling through
    messages.

    \bigskip

    \item \TF{D} Two bhikkhus are walking in the countryside, the junior bhikkhu
    is using a walking-stick. He keeps asking one question after another since
    the seniour bhikkhu doesn't seem to respond much.

    \bigskip

    \item \TF{N} Two people are offering a bhikkhu alms at the same time. The
      first person gives many things and fills his bowl, while the second person
      is waiting. The bhikkhu starts exchanging items with the second person to
      fit more nutritious items into his bowl.

    \bigskip

    \item \TF{N} While eating, a bhikkhu takes a large green pepper. He almost
    can't swallow it as a whole, but he manages to do it by pushing on it with
    his spoon.

    \bigskip

    \item \TF{N} A lay person wraps up his leftover food from the previous day
      (dry bread, soggy potatoes, mixed up rice) and offers it to a bhikkhu. He
      declines the offer, hoping to get something better later.

    \bigskip

    \item \TF{N} A bhikkhu is walking through the Dhamma Hall while a lay person
    is sitting there. Seeing the bhikkhu, they ask a few questions about
    meditation. The bhikkhu stops to stand and respond in careful detail.

    \bigskip

    \item \TF{D} A bhikkhu has been standing in front of a shop for quite a while.
    He hasn't received much food, but he leaves nonetheless.

    \bigskip

    \item \TF{N} A bhikkhu eats his alms-food in the public park. When he is
      finished, he has left-overs in his bowl but he can't see a bin, so he
      dumps it on the grass instead.

    \bigskip

    \item \TF{N} Looking into a woman's eyes while receiving alms-food.

    \bigskip

    \item \TF{D} Counting the mouthfuls while eating.

    \begin{solution}
      Counting the mouthfuls is a practice which
      (1) allows one to practice not getting distracted while eating,
      (2) gauge how much food one needs, put exactly the necessary amount in the bowl
      and leave an empty bowl when finished.
    \end{solution}

    \bigskip

    \item \TF{N} The abbot is standing up to leave, and quickly asks a question
      while the bhikkhu is chewing a mouthful. He makes sure to reply quickly
      before the abbot leaves.

    \end{parts}

  \end{problem*}

\end{exam}
