\chapter{24. Sekhiyas 2}
\renewcommand*{\theChapterTitle}{24. Sekhiyas 2}

\begin{exam}{\autoExamName}

  \begin{problem*}
    \textbf{Do} or do \textbf{Not}?

    \bigskip

    \begin{parts}

    \item \TF{D} While eating, a bhikkhu asks for water. Someone hands over a
      water bottle, but he doesn't receive it until he washed his hands.

    \bigskip

    \item \TF{N} Receiving alms-food, a bhikkhu asks a man to replace the tuna
      in tomato with tuna in oil.

    \bigskip

    \item \TF{N} Two people are offering a bhikkhu alms at the same time. The
      first person gives many things and fills his bowl, while the second person
      is waiting. The bhikkhu starts exchanging items with the second person to
      fit more nutritious items into his bowl.

    \bigskip

    \item \TF{D} While eating, collecting smaller amounts of food on the spoon.

    \bigskip

    \item \TF{N} A lay person wraps up his leftover food from the previous day
      (dry bread, soggy potatoes, mixed up rice) and offers it to a bhikkhu. He
      declines the offer, hoping to get something better later.

    \bigskip

    \item \TF{D} A bhikkhu didn't receive very nutritious food items, but he
      stood in front of the shop for quite a while now, so he leaves.

    \bigskip

    \item \TF{N} A bhikkhu eats his alms-food in the public park. When he is
      finished, he has left-overs in his bowl but he can't see a bin, so he
      dumps it on the grass instead.

    \bigskip

    \item \TF{N} Looking into a women's eyes while receiving alms-food.

    \bigskip

    \item \TF{D} Counting the mouthfuls while eating.

    \bigskip

    \item \TF{N} The abbot is standing up to leave, and quickly asks a question
      while the bhikkhu is chewing a mouthful. He makes sure to reply quickly
      before the abbot leaves.

    \end{parts}

  \end{problem*}

\end{exam}
