\chapter{Women 1}

\begin{itemize}
\tightlist
\item
  \textbf{Sg 5,} Conveying romantic messages
\item
  \textbf{Pc 6,} Lying down with a woman
\item
  \textbf{Pc 44,} Private secluded place
\item
  \textbf{Pc 45,} Unsecluded but private place
\item
  \textbf{Pc 67,} Travelling by arrangement with a woman
\end{itemize}

\section{Sg 5, Conveying romantic messages}

Only two factors: effort and object.

\textbf{Effort:} `Conveying' messages for any romantic purpose from a
momentary date to a wedding. Not business meetings.

Three stages:

\begin{itemize}
\tightlist
\item
  \emph{accepting} the request to convey a message
\item
  \emph{inquiring} at the second party
\item
  \emph{reporting} the response
\end{itemize}

Dukkata for any single stage, thullacaya for any two, sanghadisesa for
all three.

Carrying a letter without knowing the content doesn't fulfill effort.

\textbf{Object:} A man and a women who are not married to each other,
even if dealing with them via other people.

Reconciling a still married couple is not an offense. Reconciling a
divorced couple is sanghadisesa.

\textbf{Non-offenses:} messages about non-romantic errands,
e.g.~community business, a shrine, a sick person.

\section{Pc 6, Lying down with a woman}

\textbf{Object:} Female human being, even a baby, one's relative or not.

\textbf{Effort:} in the instant one lies down in the same dwelling when
a woman is lying down.

Same dwelling: one ``enclosure''. Technically the same walls and roof,
but one may consider variations (private hospital rooms).

Intention is not a factor, pacittiya even if the bhikkhu doesn't know
about the woman.

Purpose: to avoid situations where people might think that one may have
commited serious offenses. Other people might see the situation and
rumors would be damaging.

Non-offenses for roofed but no walls (pavilion) or walled but not roofed
(corral), but a good idea to avoid nonetheless.

\clearpage

\section{Pc 44, Private secluded place}

The bhikkhu sits with a woman at a secluded place, private to the eye
and ear, without another man present, aiming at privacy.

No offence if the woman entered the room later, and he didn't notice.

\section{Pc 45, Unsecluded but private place}

The bhikkhu sits with a woman at a private, but not secluded place, such
as an empty park, without another \emph{person} present.

\section{Pc 67, Travelling by arrangement with a woman}

Origin: a woman hears that a monk is going to a village and goes with
him. Later, the woman's husband heard about it and gave him a beating.

Purpose: to avoid people assuming the bhikkhu having an affair with the
woman.

\textbf{Object:} Any woman who knows what is lewd.

\textbf{Perception} is not a factor.

\textbf{Effort:}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  having made an arrangement to travel together
\item
  they travel as arranged

  \begin{itemize}
  \tightlist
  \item
    time frame as arranged
  \item
    route or place of departure doesn't count
  \end{itemize}
\item
  from one village to another (half-yojana, 8km)
\end{enumerate}

\textbf{Making an arrangement:} both gives verbal or written assent to
the arrangement.

Giving assent in silence is not an offense.

\begin{itemize}
\tightlist
\item
  if the women doesn't respond: \emph{dukkata}
\item
  if the bhikkhu doesn't respond: no offense
\end{itemize}

\subsection{Non-offenses}

\begin{itemize}
\tightlist
\item
  coincidence: they happen to travel together
\item
  the woman proposes the arrangement, and the bhikkhu doesn't give
  \emph{verbal} assent
\item
  leaving at a significantly different time than as arranged
\item
  there are dangers
\end{itemize}

\subsection{Cases}

\begin{itemize}
\tightlist
\item
  public transport
\item
  private transport (Pc 44)
\end{itemize}

