\chapter{Bowls}

\begin{itemize}
\tightlist
\item
  \textbf{NP 21,} Keeping extra bowl
\item
  \textbf{NP 22,} Asking for new bowl
\item
  \textbf{Pc 60,} Hiding another's requisites
\item
  \textbf{Pc 86,} Needle box
\end{itemize}

\section{NP 21, Keeping extra bowl}

\textbf{Object:} A suitable alms-bowl.

Five factors of an allowable bowl: proper material, size, fully paid
for, fired, not damaged beyond repair.

Allowed materials: clay, iron. Stainless steel by extension.

Forbidden materials: wood, gold, silver, pearl, beryl, crystal, bronze,
glass, tin, lead, copper. Aluminium by extension.

Proper size: Smaller than a human skull is too small. Medium size is
\textasciitilde22.5 cm diameter.

\textbf{Effort:} When the new bowl reaches one's hands, one may only
keep it as `extra' for 10 days.

One must either determine it for use, place it under shared ownership
(\emph{vikappana}), abandon it or give it away.

The offence occurs on the 10th dawnrise after receving it.

\textbf{Perception} is not a factor.

There is no offence if the bowl is lost, destroyed or stolen before the
10 days are up.

\section{NP 22, Asking for new bowl}

\emph{Origin:} a potter offers to make bowls for the bhikkhus. Some of
them abuse the offer without moderation, and the potter no longer has
time for his business work.

A bowl should be repaired as long as possible.

The offence is \emph{dukkaṭa} when asking, a \emph{nissaggiya pācittiya}
when receiving it.

`Asking' refers to asking from lay supporters. One may ask the Sangha
for a new bowl from the stores, and the community will decide whether
the request is justified.

\emph{Note:} describe the forfeiture and bowl exchange procedure.

\subsection{Non-offences}

Asking from

\begin{itemize}
\tightlist
\item
  a relative
\item
  if one was invited to ask
\item
  a new bowl with one's own resources
\item
  asking for the sake of another from relatives or through invitation
\end{itemize}

\clearpage

\section{Pc 60, Hiding another's requisites}

\textbf{Object:} bowl, robe, sitting cloth, needle box, or belt.

Hiding other requisites is a \emph{dukkaṭa} offence.

Hiding the requisites of a samanera or anagarika is a \emph{dukkaṭa}
offence.

\textbf{Perception} of whose requisite it is, is not a mitigating
factor.

A `friendly game' fulfils \textbf{Intention} all the same.

\subsection{Non-offences}

\begin{itemize}
\tightlist
\item
  putting away items to their proper place
\item
  putting it away as a teaching lesson, with the intention to give it
  back after an admonition about not leaving requisites scattered around
\end{itemize}

\section{Pc 86, Needle box}

\emph{Origin:} similar to NP 22, but with an ivory-worker. Remember to
not over-burden lay supporters by requesting items which are difficult
to make or obtain.

Forbidden materials: bone, ivory, horn.

Allowed materials for a \emph{needle box} are not explicit, but a
\emph{needle tube} is allowed of reed, bamboo, wood, lac (resin), fruit
shells, copper, conch-shell.

Receiving and using the improper item when it was requested by oneself
(e.g. given as a surprise gift) is a \emph{dukkaṭa} offence.

The intention of the rule was (a) to stop the fashionable fad, and (b)
to not be burdensome for lay people who offered their help.

The materials themselves are not banned, a number of other items are
allowed to be made of bone, ivory or horn.

A bhikkhu should be aware of the civil law regarding the materials, such
as a ban on ivory trade, or the border regulations on horn (and plant)
materials in New Zealand.

