\chapter{Questions}

\section{Introduction}

How can the monks determine if a modern item (e.g.~credit cards, sun
glasses) are allowable or not?

How does one determine whether there is full offence of a rule?

Could the abbot of a small monastery ordain a \emph{bhikkhu} or
\emph{samanera}? What should he do?

Advice on restoring one's faith after breaking a rule or having done
something unbecoming.

\section{Killing and harming}

The loved family dog of a lay supporter is very ill, and treatment will
be expensive. He asks a monk whether they should ask the vet to
euthanise the dog, or apply for treatment.

A woman asks a monk if she should get an abortion. What should the monk
say?

A monk discovers a tick on his arm. What should he do?

A monk hits an anagarika. What should the anagarika do?

A monk is attacked on the street. He pushes the attacker away and runs.
The attacker falls on the pavement and cracks his head.

A monk is attacked on the street. He is enraged and starts punching the
attacker until he goes limp and stops moving.

\section{Stealing}

A monk sneaks into the kitchen and eats an apple. Did he steal it?

A lay supporter brings an expensive sweet and gives it to a monk, saying
`I brought this for the abbot'. The monk eats a bit from it before
giving it to the abbot. Did he steal it?

A monk is visiting a monastery and makes a long phone call. The call
costs 100 EUR. The resident monks discover it on the bill and ask if
anyone knows about this call. He remains silent.

How is it possible for a monk to steal from the Sangha?

A monk drives away with the monastery car and never comes back.
Consequences?

\section{Lustful conduct}

A monk curses with lewd words in front of a woman.

A married couple wants to ask a monk how to live together peacefully.

A monk is carrying a table with a woman and he playfully pushes it into
her.

A monk receives treatment on his tooth from a female dentist.

A monk is trying on shoes in a shop. A female assistant helps to put on
a shoe and she asks, ``Is that comfortable?''

\section{Women 1}

\enlargethispage{3\baselineskip}

A monk is travelling by train, sitting in a compartment alone. At one of
the stops a woman enters and takes a seat in the compartment.

A monk stays at his parents' house for a night.

A monk is travelling by bus to visit a friend. He arrives at the bus
station, and the girlfriend of his friend is there with a car to pick
him up. She asks the monk to call his friend and tell him she will be
back at their apartment shortly.

\section{Robes 1}

A monk discovers that the seams of his cotton jacket under the arm-pit
where the cloth was joined, have come apart. What should he do?

Supporters wish to offer robe-cloth to the Community. They bring a
sample, which is a white nylon material. Is there an offence in asking
them to offer a better material?

A monk receives a nice leather-belt from a friend. Is it allowable?

A monk embroiders the sign of the Eye of Horus on his meditation
blanket. Is it allowable?

A monk takes some cloth from the stores to his kuti to make a sitting
cloth. He forgets about it for a few weeks. Is there an offence?

A monk is visiting home. His old friends invite him to the skate park.
He puts on a pair of jeans and a black T-shirt to go and see if he can
still do an ollie. Is there an offence?

A monk is chosen by the community to receive the Kathina-robe. What are
the eight Kathina duties? What is procedure when receiving the Kathina
robe? What are the Kathina privileges?

A monk is travelling by plane. He packs his sanghati in the hold
luggage. After landing, his hold luggage is missing. He registers the
missing luggage with the airport services, but has to leave without it.
The airport delivers his luggage in a few days. What are his duties?

\section{Attainments}

A monk says he can walk on water, but can't show it now because he would
lose the ability. Is there an offence? Is he still a monk?

A monk says he has a lot of metta but is still working on upekkha. Is
there an offence?

A lay supporter invites the monks for a meal: ``May the venerable
arahants come to my house tomorrow for a meal offering.'' Next day, a
few of the monks go to his house to receive the offering. Is there an
offence?

A monk is seriously ill and a group of monks visit him. He says that he
is not afraid of death. Is there an offence?

\section{Robes 2}

A monk takes a tea-towel from the kitchen to his kuti. Should he bindu
and determine it?

A monk takes bits of left-over cloth from the sewing room and makes a
belt-pouch. Should he bindu and determine it?

\section{Food 1}

What is miso and why is it life-time?

Is rice- or almod milk allowable in the afternoon?

\section{Food 2}

What is the lifetime of the following items?

\begin{itemize}
\tightlist
\item
  Fruit juice in tetra-pack
\item
  Unsweetened soya milk
\item
  Margarine (from veg. oil)
\item
  Butter (dairy)
\item
  Fried onions
\item
  Coca-Cola
\item
  Cheese with red pepper spicing
\item
  Cheese with onion pieces
\item
  Coffee-mate powder
\item
  Carrot juice
\item
  Jelly
\end{itemize}

